Section Two

### 2020-12-01 new2020/sections/sec02.Rmd
fix missing cite

An important aspect missing from the setup, however, is information about *physical interactions* between separate routes (which may share common roads) which is critical to accurate prediction @citep:Jeong_2005,;. @citet:Elliott_2020; describe a simple process for combining route information from static GTFS data to obtain a *transit road network*, expressing each route as a sequence of physical roads. The benefit of this is that vehicles servicing different routes but using the same road(s) can share information about their speed. This greatly enhances the amount of data available for forecasting arrival times.

@@

Estimation of recursive Bayesian models takes many forms, most notably the Kalman filter @citep:; which has been widely used in transit modelling @citep:Wall_1999,Dailey_2001,Cathey_2003,Jeong_2005;. However, as discussed by @citet:Hans_2015,Elliott_2020;, it makes several assumptions which do not work so well with transit vehicles, namely that the errors are normal with a single mode. The particle filter @citep:Gordon_1993; provides a numerical approach to solving recursive Bayesian models, much the same as Markov chain Monte Carlo provides estimates of standard posterior distributions. In a particle filter, the state $\vec x_k$ is represented by a sample of $N$\ particles $\{\vec x_k\vi\}_{i=1}^N$ with associated weights $\{w_k\vi\}_{i=1}^N$. We use the delta measure $\delta$ to express the approximated state
\[
   p(\vec x_{k-1} | \vec y_{0:k-1}) \approx \sum_{i=1}^N w_{k-1}^{(i)} \DiracMeasure{\vec x_{k-1}^{(i)}}{\vec x_{k-1}}.
\]

>>>>

An important aspect missing from the setup, however, is information about *physical interactions* between separate routes (which may share common roads) which is critical to accurate prediction @citep:Jeong_2005;. @citet:Elliott_2020; describe a simple process for combining route information from static GTFS data to obtain a *transit road network*, expressing each route as a sequence of physical roads. The benefit of this is that vehicles servicing different routes but using the same road(s) can share information about their speed. This greatly enhances the amount of data available for forecasting arrival times.

@@

Estimation of recursive Bayesian models takes many forms, most notably the Kalman filter @citep:kf; which has been widely used in transit modelling @citep:Wall_1999,Dailey_2001,Cathey_2003,Jeong_2005;. However, as discussed by @citet:Hans_2015,Elliott_2020;, it makes several assumptions which do not work so well with transit vehicles, namely that the errors are normal with a single mode. The particle filter @citep:Gordon_1993; provides a numerical approach to solving recursive Bayesian models, much the same as Markov chain Monte Carlo provides estimates of standard posterior distributions. In a particle filter, the state $\vec x_k$ is represented by a sample of $N$\ particles $\{\vec x_k\vi\}_{i=1}^N$ with associated weights $\{w_k\vi\}_{i=1}^N$. We use the delta measure $\delta$ to express the approximated state
\[
   p(\vec x_{k-1} | \vec y_{0:k-1}) \approx \sum_{i=1}^N w_{k-1}^{(i)} \DiracMeasure{\vec x_{k-1}^{(i)}}{\vec x_{k-1}}.
\]
